```{r include=FALSE, cache=FALSE}
# DO NOT REMOVE
# THIS IS FOR SETTING SOME PLOTTING PARAMETERS SO THAT YOUR PLOTS DON'T TAKE UP TOO MUCH SPACE
# IF YOU WOULD LIKE TO CHANGE THESE, SEE HELP FILES ON THE par() FUNCTION
# OR ASK FOR HELP
library(knitr)
## set global chunk options
opts_chunk$set(fig.path='figure/manual-', cache.path='cache/manual-', fig.align='center', fig.show='hold', par=TRUE)
## tune details of base graphics (http://yihui.name/knitr/hooks)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.2,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
})
```

```{r echo=FALSE}
# load data



```


### Part 1 - Introduction:


The goal of this research is to find if there is a link between Pollution levels and incidence of Tuberculosis. There have been studies that indicate that there is a relation between indoor pollution / smoking to TB. However, in this proposal I am looking to study if PM2.5 Air Pollution has any effect on incidence of TB.

This is an observational study



### Part 2 - Data:

The data is sourced from the data respository maintained by World Bank. Hence it is secondary data. This secondary data is collected from the World Bank website.

World Bank in turn gets the TB data from the Global Tuberculosis Report of World Health Organization (WHO)

similarly, the pollution data is sourced by World Bank from Brauer, M. et al. 2015. "Ambient Air Pollution Exposure Estimation for the Global Burden of Disease 2013." (Paper submitted for publication.), Institute for Health Metrics and Evaluation, University of Washington, Seattle.


The data is a part of the World Bank data. The required data for the 2 variables in question are available for download from the website. 

http://data.worldbank.org/indicator/EN.ATM.PM25.MC.ZS?page=5&display=default

http://data.worldbank.org/indicator/SH.TBS.INCD?display=default


The Response Variable is Incidence of TB

In this study, I am trying to explain an increase / decrease in the occurance of TB based on the increase / decrease in the pollution rate. The Response variable is Incidence of TB. 

It is a numeric variable.

Incidence of tuberculosis (per 100,000 people) is the estimated number of new and relapse tuberculosis cases arising in a given year, expressed as the rate per 100,000 population. All forms of TB are included, including cases in people living with HIV. 



The Explanatory variable is the Pollution data

PM2.5 air pollution, population exposed to levels exceeding WHO guideline value (% of total) is defined as the portion of a country's population living in places where mean annual concentrations of PM2.5 are greater than 10 micrograms per cubic meter, the guideline value recommended by the World Health Organization as the lower end of the range of concentrations over which adverse health effects due to PM2.5 exposure have been observed.


generate the proportion increase




### Part 3 - Exploratory data analysis:




### Part 4 - Inference:


### Part 5 - Conclusion: 


### References:


### Appendix (optional):

Remove this section if you don't have an appendix
